---
title: "BigData Project"
author: "Saksham Dixit"
date: "April 23, 2017"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(corrplot)
library(fBasics)
library(caret)
library(penalized)
library(glmnet)
library(e1071)
library(rpart)
library(kernlab)
library(nnet)
library(DMwR)
library(ROCR)
library(ROSE)
```
# Data Exploration

Using a 25% sample from the training data for exploration and feature selection

```{r}
smpTrain = read.csv(file ='smpTrain.csv',header= FALSE,stringsAsFactors = TRUE)

#assigning predictor names
colnames(smpTrain) <- c('age','workclass','fnlwgt','education','education_num',
                        'marital_status', 'occupation', 'relationship', 'race',
                        'sex', 'capital_gain', 'capital_loss', 'hpw',
                        'native_country','salarymorethan50k')

smpTrain$age <- as.numeric(smpTrain$age)
smpTrain$fnlwgt <- as.numeric(smpTrain$fnlwgt)
smpTrain$capital_gain <- as.numeric(smpTrain$capital_gain)
smpTrain$capital_loss <- as.numeric(smpTrain$capital_loss)
smpTrain$hpw <- as.numeric(smpTrain$hpw)

#removing education_num as it is redundant
smpTrain <- smpTrain[-5]
#Seeing a summary of the sample data
summary(smpTrain)

print(c('frequency of the target',table(smpTrain$salarymorethan50k)),quote= FALSE)

#compute the skewdness of numeric predictors
predictors <- smpTrain[c(1,3,10,11,12)]
skewValues <- apply(predictors,2,skewness)
print(skewValues)
```
# Loading the training data and test data

```{r}
train1 <- read.csv(file = 'train1.csv',header = FALSE, stringsAsFactors = TRUE)

#assigning predictor names
colnames(train1) <- c('age','workclass','fnlwgt','education','marital_status',
                      'occupation', 'relationship','race', 'sex','capital_gain',
                      'capital_loss','hpw','native_country','salarymorethan50k')

train1$age  <- as.numeric(train1$age)
train1$fnlwgt  <- as.numeric(train1$fnlwgt)
train1$capital_gain  <- as.numeric(train1$capital_gain)
train1$capital_loss  <- as.numeric(train1$capital_loss)
train1$hpw  <- as.numeric(train1$hpw)
train1$salarymorethan50k <- as.factor(train1$salarymorethan50k)

levels(train1$occupation) <- c('Tech-support', 'Craft-repair', 'Other-service', 'Sales', 'Exec-managerial', 'Prof-specialty', 'Handlers-cleaners', 'Machine-op-inspct', 'Adm-clerical', 'Farming-fishing', 'Transport-moving', 'Priv-house-serv', 'Protective-serv', 'Armed-Forces')

levels(train1$native_country) <- c('United-States', 'Cambodia', 'England', 'Puerto-Rico', 'Canada', 'Germany', 'Outlying-US(Guam-USVI-etc)', 'India', 'Japan', 'Greece', 'South', 'China', 'Cuba', 'Iran', 'Honduras', 'Philippines', 'Italy', 'Poland', 'Jamaica', 'Vietnam', 'Mexico', 'Portugal', 'Ireland', 'France', 'Dominican-Republic', 'Laos', 'Ecuador', 'Taiwan', 'Haiti', 'Columbia', 'Hungary', 'Guatemala', 'Nicaragua', 'Scotland', 'Thailand', 'Yugoslavia', 'El-Salvador', 'Trinadad&Tobago', 'Peru', 'Hong', 'Holand-Netherlands')

train2 <- read.csv(file = 'train2.csv',header = FALSE, stringsAsFactors = TRUE)

#assigning predictor names
colnames(train2) <- c('age','workclass','fnlwgt','education','marital_status',
                      'occupation', 'relationship','race', 'sex','capital_gain',
                      'capital_loss','hpw','native_country','salarymorethan50k')

train2$age  <- as.numeric(train2$age)
train2$fnlwgt  <- as.numeric(train2$fnlwgt)
train2$capital_gain  <- as.numeric(train2$capital_gain)
train2$capital_loss  <- as.numeric(train2$capital_loss)
train2$hpw  <- as.numeric(train2$hpw)
train2$salarymorethan50k <- as.factor(train2$salarymorethan50k)


levels(train2$occupation) <- c('Tech-support', 'Craft-repair', 'Other-service', 'Sales', 'Exec-managerial', 'Prof-specialty', 'Handlers-cleaners', 'Machine-op-inspct', 'Adm-clerical', 'Farming-fishing', 'Transport-moving', 'Priv-house-serv', 'Protective-serv', 'Armed-Forces')

levels(train2$native_country) <- c('United-States', 'Cambodia', 'England', 'Puerto-Rico', 'Canada', 'Germany', 'Outlying-US(Guam-USVI-etc)', 'India', 'Japan', 'Greece', 'South', 'China', 'Cuba', 'Iran', 'Honduras', 'Philippines', 'Italy', 'Poland', 'Jamaica', 'Vietnam', 'Mexico', 'Portugal', 'Ireland', 'France', 'Dominican-Republic', 'Laos', 'Ecuador', 'Taiwan', 'Haiti', 'Columbia', 'Hungary', 'Guatemala', 'Nicaragua', 'Scotland', 'Thailand', 'Yugoslavia', 'El-Salvador', 'Trinadad&Tobago', 'Peru', 'Hong', 'Holand-Netherlands')

train3 <- read.csv(file = 'train3.csv',header = FALSE, stringsAsFactors = TRUE)

#assigning predictor names
colnames(train3) <- c('age','workclass','fnlwgt','education','marital_status',
                      'occupation', 'relationship','race', 'sex','capital_gain',
                      'capital_loss','hpw','native_country','salarymorethan50k')

train3$age  <- as.numeric(train3$age)
train3$fnlwgt  <- as.numeric(train3$fnlwgt)
train3$capital_gain  <- as.numeric(train3$capital_gain)
train3$capital_loss  <- as.numeric(train3$capital_loss)
train3$hpw  <- as.numeric(train3$hpw)
train3$salarymorethan50k <- as.factor(train3$salarymorethan50k)


levels(train3$occupation) <- c('Tech-support', 'Craft-repair', 'Other-service', 'Sales', 'Exec-managerial', 'Prof-specialty', 'Handlers-cleaners', 'Machine-op-inspct', 'Adm-clerical', 'Farming-fishing', 'Transport-moving', 'Priv-house-serv', 'Protective-serv', 'Armed-Forces')

levels(train3$native_country) <- c('United-States', 'Cambodia', 'England', 'Puerto-Rico', 'Canada', 'Germany', 'Outlying-US(Guam-USVI-etc)', 'India', 'Japan', 'Greece', 'South', 'China', 'Cuba', 'Iran', 'Honduras', 'Philippines', 'Italy', 'Poland', 'Jamaica', 'Vietnam', 'Mexico', 'Portugal', 'Ireland', 'France', 'Dominican-Republic', 'Laos', 'Ecuador', 'Taiwan', 'Haiti', 'Columbia', 'Hungary', 'Guatemala', 'Nicaragua', 'Scotland', 'Thailand', 'Yugoslavia', 'El-Salvador', 'Trinadad&Tobago', 'Peru', 'Hong', 'Holand-Netherlands')


#load test data
test <- read.csv(file = 'testData.csv',header = FALSE, stringsAsFactors = TRUE)

#assigning predictor names
colnames(test) <- c('age','workclass','fnlwgt','education','marital_status',
                      'occupation', 'relationship','race', 'sex','capital_gain',
                      'capital_loss','hpw','native_country','salarymorethan50k')

test$age  <- as.numeric(test$age)
test$fnlwgt  <- as.numeric(test$fnlwgt)
test$capital_gain  <- as.numeric(test$capital_gain)
test$capital_loss  <- as.numeric(test$capital_loss)
test$hpw  <- as.numeric(test$hpw)
test$salarymorethan50k <- as.factor(test$salarymorethan50k)


levels(test$occupation) <- c('Tech-support', 'Craft-repair', 'Other-service', 'Sales', 'Exec-managerial', 'Prof-specialty', 'Handlers-cleaners', 'Machine-op-inspct', 'Adm-clerical', 'Farming-fishing', 'Transport-moving', 'Priv-house-serv', 'Protective-serv', 'Armed-Forces')

levels(test$native_country) <- c('United-States', 'Cambodia', 'England', 'Puerto-Rico', 'Canada', 'Germany', 'Outlying-US(Guam-USVI-etc)', 'India', 'Japan', 'Greece', 'South', 'China', 'Cuba', 'Iran', 'Honduras', 'Philippines', 'Italy', 'Poland', 'Jamaica', 'Vietnam', 'Mexico', 'Portugal', 'Ireland', 'France', 'Dominican-Republic', 'Laos', 'Ecuador', 'Taiwan', 'Haiti', 'Columbia', 'Hungary', 'Guatemala', 'Nicaragua', 'Scotland', 'Thailand', 'Yugoslavia', 'El-Salvador', 'Trinadad&Tobago', 'Peru', 'Hong', 'Holand-Netherlands')

```
#Seeing the distribution of the classes in the three sets
```{r}
table(train1$salarymorethan50k)
table(train2$salarymorethan50k)
table(train3$salarymorethan50k)
table(test$salarymorethan50k)
```
#sMOTE the data to balance the classes
```{r}
set.seed(111)
smoteTrain1 <- SMOTE(salarymorethan50k ~ .,train1,perc.over = 100, perc.under = 200)
table(smoteTrain1$salarymorethan50k)


levels(smoteTrain1$occupation) <- c('Tech-support', 'Craft-repair', 'Other-service', 'Sales', 'Exec-managerial', 'Prof-specialty', 'Handlers-cleaners', 'Machine-op-inspct', 'Adm-clerical', 'Farming-fishing', 'Transport-moving', 'Priv-house-serv', 'Protective-serv', 'Armed-Forces')

levels(smoteTrain1$native_country) <- c('United-States', 'Cambodia', 'England', 'Puerto-Rico', 'Canada', 'Germany', 'Outlying-US(Guam-USVI-etc)', 'India', 'Japan', 'Greece', 'South', 'China', 'Cuba', 'Iran', 'Honduras', 'Philippines', 'Italy', 'Poland', 'Jamaica', 'Vietnam', 'Mexico', 'Portugal', 'Ireland', 'France', 'Dominican-Republic', 'Laos', 'Ecuador', 'Taiwan', 'Haiti', 'Columbia', 'Hungary', 'Guatemala', 'Nicaragua', 'Scotland', 'Thailand', 'Yugoslavia', 'El-Salvador', 'Trinadad&Tobago', 'Peru', 'Hong', 'Holand-Netherlands')

set.seed(111)
smoteTrain2 <- SMOTE(salarymorethan50k ~ .,train2,perc.over = 100, perc.under = 200)
table(smoteTrain2$salarymorethan50k)


levels(smoteTrain2$occupation) <- c('Tech-support', 'Craft-repair', 'Other-service', 'Sales', 'Exec-managerial', 'Prof-specialty', 'Handlers-cleaners', 'Machine-op-inspct', 'Adm-clerical', 'Farming-fishing', 'Transport-moving', 'Priv-house-serv', 'Protective-serv', 'Armed-Forces')

levels(smoteTrain2$native_country) <- c('United-States', 'Cambodia', 'England', 'Puerto-Rico', 'Canada', 'Germany', 'Outlying-US(Guam-USVI-etc)', 'India', 'Japan', 'Greece', 'South', 'China', 'Cuba', 'Iran', 'Honduras', 'Philippines', 'Italy', 'Poland', 'Jamaica', 'Vietnam', 'Mexico', 'Portugal', 'Ireland', 'France', 'Dominican-Republic', 'Laos', 'Ecuador', 'Taiwan', 'Haiti', 'Columbia', 'Hungary', 'Guatemala', 'Nicaragua', 'Scotland', 'Thailand', 'Yugoslavia', 'El-Salvador', 'Trinadad&Tobago', 'Peru', 'Hong', 'Holand-Netherlands')


set.seed(111)
smoteTrain3 <- SMOTE(salarymorethan50k ~ .,train3,perc.over = 100, perc.under = 200)
table(smoteTrain3$salarymorethan50k)


levels(smoteTrain3$occupation) <- c('Tech-support', 'Craft-repair', 'Other-service', 'Sales', 'Exec-managerial', 'Prof-specialty', 'Handlers-cleaners', 'Machine-op-inspct', 'Adm-clerical', 'Farming-fishing', 'Transport-moving', 'Priv-house-serv', 'Protective-serv', 'Armed-Forces')

levels(smoteTrain3$native_country) <- c('United-States', 'Cambodia', 'England', 'Puerto-Rico', 'Canada', 'Germany', 'Outlying-US(Guam-USVI-etc)', 'India', 'Japan', 'Greece', 'South', 'China', 'Cuba', 'Iran', 'Honduras', 'Philippines', 'Italy', 'Poland', 'Jamaica', 'Vietnam', 'Mexico', 'Portugal', 'Ireland', 'France', 'Dominican-Republic', 'Laos', 'Ecuador', 'Taiwan', 'Haiti', 'Columbia', 'Hungary', 'Guatemala', 'Nicaragua', 'Scotland', 'Thailand', 'Yugoslavia', 'El-Salvador', 'Trinadad&Tobago', 'Peru', 'Hong', 'Holand-Netherlands')


```
# function that takes the majority vote

```{r}
judge <- function(row){
        row <- as.numeric(row)
        if (sum(row) >= 2){
                pred <- 1
        }else{
                pred <- 0
        }
        return(pred)
}
```

#Building Decision Tree


```{r}
#Prune back the tree to avoid overfitting the data we select a tree size that
#minimizes the cross-validated error

createTreeModel <- function(trainData){
   treeModel <- rpart(formula = salarymorethan50k ~ .,data = trainData,
                        method = 'class',control = rpart.control(maxdepth = 30))
   
  ptreeModel <- prune(treeModel,
           cp =treeModel$cptable[which.min(treeModel$cptable[,"xerror"]),"CP"])

   
     return(ptreeModel)
}

#for first set of training Data
tree1 <- createTreeModel(smoteTrain1)
predictTree1 <- predict(tree1,test[,-14],type = 'class')
tree1Cnm <- confusionMatrix(data = predictTree1,
                               reference = test$salarymorethan50k)
print(tree1Cnm$overall[1])

#for second set of training Data
tree2 <- createTreeModel(smoteTrain2)
predictTree2 <- predict(tree2,test[,-14], type ='class')
tree2Cnm <- confusionMatrix(data = predictTree2,
                               reference = test$salarymorethan50k)
print(tree2Cnm$overall[1])

#for third set of training Data
tree3 <- createTreeModel(smoteTrain3)
predictTree3 <- predict(tree3,test[,-14], type ='class')
tree3Cnm <- confusionMatrix(data = predictTree3,
                               reference = test$salarymorethan50k)
print(tree3Cnm$overall[1])

#combine all the predictions into one 
treePredictions <- data.frame(predictTree1,predictTree2,predictTree3)

#take the majority vote
finalTreePred <- as.factor(apply(treePredictions,1,judge))

#build the confusin matrix
finalTreeCnm <- confusionMatrix(data = finalTreePred,
                               reference = test$salarymorethan50k)
#final accuracy
print(finalTreeCnm$table)
print(finalTreeCnm$overall[1])


#plot ROC
predictTree1Prob <- predict(tree1,test[,-14],type = 'prob')
predictTree2Prob <- predict(tree2,test[,-14],type = 'prob')
predictTree3Prob <- predict(tree3,test[,-14],type = 'prob')

treePredictionsProb <- data.frame(predictTree1Prob[,2],predictTree2Prob[,2],
                                  predictTree3Prob[,2])

finalTreePredProb <- apply(treePredictionsProb,1,prod)

prefTree<-performance(prediction(finalTreePredProb,test$salarymorethan50k),
                  measure = "tpr",x.measure = "fpr")
plot(prefTree)

#auc
aucTree<-performance(prediction(finalTreePredProb,test$salarymorethan50k),
                 measure = "auc")
print(c("Area Under the Curve: ", aucTree@y.values[[1]]),quote = FALSE)

```
# Use Naive Bayes 
```{r}

createNbMOdel <- function(trainData){
         
        nbModel <- naiveBayes(formula = salarymorethan50k ~ .,data = trainData)
         
        return(nbModel)
}

#create modles for the three data sets
nbModel1 <- createNbMOdel(smoteTrain1)  
nbModel2 <- createNbMOdel(smoteTrain2)
nbModel3 <- createNbMOdel(smoteTrain3)

#get the predictions of the models
predNbModel1 <- predict(nbModel1,test, type = 'class') 
predNbModel2 <- predict(nbModel2,test, type = 'class')
predNbModel3 <- predict(nbModel3,test, type = 'class')

#get the accuracy of each individual model
nbCnm1 <- confusionMatrix(data = predNbModel1,
                               reference = test$salarymorethan50k)
print(nbCnm1$overall[1])

nbCnm2 <- confusionMatrix(data = predNbModel2,
                               reference = test$salarymorethan50k)
print(nbCnm2$overall[1])
nbCnm3 <- confusionMatrix(data = predNbModel3,
                               reference = test$salarymorethan50k)
print(nbCnm3$overall[1])

#collect all the predictions
predictionsNb <- data.frame(predNbModel1,predNbModel2,
                          predNbModel3)

finalNbPred <- as.factor(apply(predictionsNb,1,judge))
finalNbCnm <- confusionMatrix(data = finalNbPred,
                               reference = test$salarymorethan50k)

#print the confusion matrix and the accuracy
print(finalNbCnm$table)
print(finalNbCnm$overall[1])

#plot ROC

predNbModel1Prob <- predict(nbModel1,test, type = 'raw') 
predNbModel2Prob <- predict(nbModel2,test, type = 'raw')
predNbModel3Prob <- predict(nbModel3,test, type = 'raw')

nbPredictionsProb <- data.frame(predNbModel1Prob[,2],predNbModel2Prob[,2],
                                predNbModel3Prob[,2])

finalNBPredProb <- apply(nbPredictionsProb,1,prod)

pref<-performance(prediction(finalNBPredProb,test$salarymorethan50k),
                  measure = "tpr",x.measure = "fpr")
plot(pref)

#auc
aucNB <- performance(prediction(finalNBPredProb,test$salarymorethan50k),measure = "auc")
print(c("Area Under the Curve: ", aucNB@y.values[[1]]),quote = FALSE)

```
# Logistic Regression Model

```{r,message = FALSE,warning = FALSE}
createLogisticModel <- function(trainData){
              
        logModel <- glm(salarymorethan50k ~ ., family= binomial(),
                        data = trainData, model = TRUE)
        return(logModel)
        
}

logModel1 <- createLogisticModel(smoteTrain1)
logModel2 <- createLogisticModel(smoteTrain2)
logModel3 <- createLogisticModel(smoteTrain3)

#making sure models have all required levels
logModel1$xlevels[["occupation"]] <- union(logModel1$xlevels[["occupation"]], 
                                           levels(test$occupation))
logModel2$xlevels[["occupation"]] <- union(logModel2$xlevels[["occupation"]], 
                                           levels(test$occupation))

logModel3$xlevels[["occupation"]] <- union(logModel3$xlevels[["occupation"]], 
                                           levels(test$occupation))
logModel1$xlevels[["native_country"]] <- union(logModel1$xlevels[["native_country"]], 
                                           levels(test$native_country))
logModel2$xlevels[["native_country"]] <- union(logModel2$xlevels[["native_country"]], 
                                           levels(test$native_country))

logModel3$xlevels[["native_country"]] <- union(logModel3$xlevels[["native_country"]], 
                                           levels(test$native_country))

#getting the predictions for each individual model
predictLogModel1Prob <- predict(logModel1,test, type = 'response')
predictLogModel2Prob <- predict(logModel2,test, type = 'response')
predictLogModel3Prob <- predict(logModel3,test, type = 'response')

#collecting the predictions 
logPredictionsProb <- data.frame(predictLogModel1Prob,
                                 predictLogModel2Prob,predictLogModel3Prob)

#getting the class form the predicted probabilities
predictLogModel1 <- as.factor(ifelse( predictLogModel1Prob >= 0.5,1,0))
predictLogModel2 <- as.factor(ifelse( predictLogModel2Prob >= 0.5,1,0))
predictLogModel3 <- as.factor(ifelse( predictLogModel3Prob >= 0.5,1,0)) 


log1Cnm <-  confusionMatrix(data = predictLogModel1,
                               reference = test$salarymorethan50k)
print(log1Cnm$overall[1])
log2Cnm <-  confusionMatrix(data = predictLogModel2,
                               reference = test$salarymorethan50k)
print(log2Cnm$overall[1])
log3Cnm <-  confusionMatrix(data = predictLogModel3,
                               reference = test$salarymorethan50k)
print(log3Cnm$overall[1])
#Combining the predictions on the three sets
logModelPredictions <- data.frame(predictLogModel1,predictLogModel2,
                                                              predictLogModel3)

finalLogPred <- as.factor(apply(logModelPredictions,1,judge))
finalLogCnm <-  confusionMatrix(data = finalLogPred,
                               reference = test$salarymorethan50k)
print(finalLogCnm$overall[1])
print(finalLogCnm$table)

#plot ROC

finalLogPredProb <- apply(logPredictionsProb,1,prod)
pref<-performance(prediction(finalLogPredProb,test$salarymorethan50k),
                  measure = "tpr",x.measure = "fpr")
plot(pref)

#auc
aucLog <- performance(prediction(finalLogPredProb,test$salarymorethan50k),measure = "auc")
print(c("Area Under the Curve: ", aucLog@y.values[[1]]),quote = FALSE)

```
# Penalized Logistic Model: Does not work

```{r}
createPenalizedModel <- function(trainData){
        penModel <- glmnet(x = trainData[,-14], y = trainData[,14],
                           family = "binomial",alpha = 1)
        return(penModel)
}

penModel1 <- createPenalizedModel(train1)
```

# Support Vector Machine

```{r}
#Function to create SVM model and return predictions
createSVM <- function(trainData){
        
        SVM <- ksvm( salarymorethan50k~.,data=trainData, type="C-bsvc",
                    kernel= 'rbfdot', C= 1.25, prob.model=TRUE)
        
          return(SVM)
}

#Create models for the three data sets
SVM1 <- createSVM(smoteTrain1)
SVM2 <- createSVM(smoteTrain2)
SVM3 <- createSVM(smoteTrain3)

#get the predictions for the three models
predictSVM1 <- predict(SVM1, test, type = 'response')
predictSVM2 <- predict(SVM2, test, type = 'response')
predictSVM3 <- predict(SVM3, test, type = 'response')

#Get the acurracy of each individual model
SVMCnm1 <- confusionMatrix(data = predictSVM1, 
                                  reference = test$salarymorethan50k)
print(SVMCnm1$overall[1])
SVMCnm2 <- confusionMatrix(data = predictSVM2, 
                                  reference = test$salarymorethan50k)
print(SVMCnm2$overall[1])
SVMCnm3 <- confusionMatrix(data = predictSVM3, 
                                  reference = test$salarymorethan50k)
print(SVMCnm3$overall[1])

#majority vote and final accuracy calculation
predictionSVM <- data.frame(predictSVM1,predictSVM2,predictSVM3)
finalSVMpredict <- as.factor(apply(predictionSVM,1,judge))

finalSVMCnm <- confusionMatrix(data = finalSVMpredict, 
                                  reference = test$salarymorethan50k)
print(finalSVMCnm$overall[1])
print(finalSVMCnm$table)

#Plot ROC
predictSVM1Prob <- predict(SVM1, test, type = 'probabilities')
predictSVM2Prob <- predict(SVM1, test, type = 'probabilities')
predictSVM3Prob <- predict(SVM1, test, type = 'probabilities')

SVMPredictionsProb <- data.frame(predictSVM1Prob[,2],predictSVM2Prob[,2],
                                                    predictSVM3Prob[,2])

finalSVMPredProb <- apply(SVMPredictionsProb,1,prod)
finalSVMPredProb <- finalSVMPredProb

prefSVM <-performance(prediction(finalSVMPredProb,test$salarymorethan50k),
                  measure = "tpr",x.measure = "fpr")
plot(prefSVM)

#auc
aucSVM <- performance(prediction(finalSVMPredProb,test$salarymorethan50k),measure = "auc")
print(c("Area Under the Curve: ", aucSVM@y.values[[1]]),quote = FALSE)

```
## Artificial Neural Network 
```{r,message = FALSE,warning = FALSE}
createANN <- function(trainData,testData,nodes,tune = FALSE){
     
        ANNModel <- nnet(salarymorethan50k ~., data = trainData, size = nodes,
                         maxit = 2000,trace = FALSE, MaxNWts = 20000)
        
        predANN <- predict(ANNModel,testData[,-14], type = 'class')
        ANNCnm <- confusionMatrix(data = predANN, 
                                  reference = testData$salarymorethan50k)
        if(tune == TRUE){
                return(ANNCnm$overall[1])
        }else if (tune == FALSE){
                return(ANNModel)
        }        
}

tuneANN <- function(trainData){
        
        #Using 5-fold CrossValiadation on 1st training set
        set.seed(111)
        cv_train_index <- createFolds(seq_len(nrow(trainData)), k = 5, list = TRUE,
                                                             returnTrain = TRUE)
        acc <- NA          
        avgacc <- NA
        k = 1
        for(j in seq(from = 1,to = 10 ,by = 1)){
                acc <- NA
                for(i in 1:5){
                        acc[i] <- createANN(trainData[cv_train_index[[i]],],
                                           trainData[-cv_train_index[[i]],], j,tune = TRUE)
                 }
        avgacc[k] <- mean(acc)
        k <- k+1
        }
       
ledger_mat <- data.frame(seq(from = 1, to = 10, by = 1),avgacc)

tunePlot <- ggplot(ledger_mat, aes(x = ledger_mat[,1], y = ledger_mat[,2]))+
        geom_point()+ geom_line()+xlab('No of Nodes')+ylab('Accuracy')+
        scale_x_continuous(breaks = round(seq(1,10,1),1))

print(tunePlot)
return(ledger_mat[which.max(ledger_mat[,2]),1])
}

#tune first training set using 5 fold cross validation
size1 <- tuneANN(smoteTrain1)
print(c('No of nodes in the hidden layer',size1), quote = FALSE)
#tune second training set using 5 fold cross validation
size2 <- tuneANN(smoteTrain2)
print(c('No of nodes in the hidden layer',size2), quote = FALSE)
#tune third training set using 5 fold cross validation
size3 <- tuneANN(smoteTrain3)
print(c('No of nodes in the hidden layer',size3), quote = FALSE)

#creating the three models using the tuned hyper parameter
ANN1 <- createANN(smoteTrain1,test,nodes = size1)
ANN2 <- createANN(smoteTrain2,test,nodes = size2)
ANN3 <- createANN(smoteTrain3,test,nodes = size3)

predANN1 <- predict(ANN1,test, type = 'class')
ANNCnm1 <- confusionMatrix(data = predANN1, 
                                  reference = test$salarymorethan50k)
print(ANNCnm1$table)
print(ANNCnm1$overall[1])

predANN2 <- predict(ANN2,test, type = 'class')
ANNCnm2 <- confusionMatrix(data = predANN2, 
                                  reference = test$salarymorethan50k)
print(ANNCnm2$table)
print(ANNCnm2$overall[1])

predANN3 <- predict(ANN3,test, type = 'class')
ANNCnm3 <- confusionMatrix(data = predANN3, 
                                  reference = test$salarymorethan50k)
print(ANNCnm3$table)
print(ANNCnm3$overall[1])


#collection all the 3 sets of predicitions
predictionANN <- data.frame(predANN1,predANN2,predANN3)

#taking a majority vote for each record of test data
finalANNpredict <- as.factor(apply(predictionANN,1,judge))

finalANNCnm <- confusionMatrix(data = finalANNpredict, 
                                  reference = test$salarymorethan50k)
print(finalANNCnm$overall[1])
print(finalANNCnm$table)


#Pot ROC
predANN1Prob <- predict(ANN1,test, type = 'raw')
predANN2Prob <- predict(ANN2,test, type = 'raw')
predANN3Prob <- predict(ANN3,test, type = 'raw')

ANNPredictionsProb <- data.frame(predANN1Prob,predANN2Prob,predANN3Prob)

finalANNPredProb <- apply(ANNPredictionsProb,1,prod)

prefANN<-performance(prediction(finalANNPredProb,test$salarymorethan50k),
                  measure = "tpr",x.measure = "fpr")
plot(prefANN)

#auc
aucANN <- performance(prediction(finalANNPredProb,test$salarymorethan50k),measure = "auc")
print(c("Area Under the Curve: ", aucANN@y.values[[1]]),quote = FALSE)

        
```

