---
title: "Final_Project"
author: "Saksham Dixit"
date: "May 3, 2017"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(C50)
library(glmnet)
library(caret)
library(DMwR)
library(xgboost)
library(ROCR)
library(covTest)
library(penalized)
```
#PART 1
```{r}

# This part conatins data exploration, calculations and graphs

data(churn)
train <- churnTrain
test <- churnTest

churnTot <- rbind(churnTrain,churnTest)
churnYes <- churnTot[churnTot$churn == 'yes',]
churnNo <-  churnTot[churnTot$churn == 'no',]

churnRate <- (length(churnTot[churnTot$churn == 'yes',20])
                                /length(churnTot$churn))
              
df <- data.frame(c(length(churnYes$churn),length(churnNo$churn)),
                 c('Churn','Continuing'))
colnames(df) <- c('number','class')        

#pie chart of customer churn

slices <- df$number 
lbls <- df$class
prct <- round(slices/sum(slices)*100)
lbls <- paste(lbls, prct) 
lbls <- paste(lbls,"%",sep="")

acLengthNO <- mean(churnNo$account_length)
acLengthYes <- mean(churnYes$account_length)

# percentage churn by state

states <- levels(churnYes$state)
yesCount <- NA
noCount <- NA
churnPerc <- NA
for(i in seq_len(length(states))){
        yes <- NA
        no <- NA
        
        yes <- ifelse (churnYes$state == states[i],1,0)
        no <- ifelse (churnNo$state == states[i],1,0)
        yesCount[i] <- sum(yes)
        noCount[i] <- sum(no)
        churnPerc[i] <- yesCount[i]/(sum(yesCount[i],noCount[i]))
}


churnState <- data.frame(states,churnPerc)
colnames(churnState) <- c('State','Churn_Perc')

gChurnstatePerc <- ggplot(churnState,aes(churnState$State,churnState$Churn_Perc))+
        geom_bar(stat = 'identity',size = churnState$Churn_Perc)+
        labs(title ='Percentage Churn By State', x = 'State', y = 'Percentage Churn')




# percentage of customer churning with international plan

intPlanCont <- ifelse(churnNo$international_plan == 'yes',1,0)
intPlanContPerc <- sum(intPlanCont)/length(churnTot$international_plan)

intPlanChurn <- ifelse(churnYes$international_plan == 'yes',1,0)
intPlanChurnPerc <- sum(intPlanChurn)/length(churnTot$international_plan)

intPlanTot <- ifelse(churnTot$international_plan == 'yes',1,0)
intPlanPerc <- sum(intPlanTot)/length(churnTot$international_plan)

intPlanDf <- data.frame(c(intPlanPerc,intPlanContPerc,intPlanChurnPerc),
                        c('Total % customers With International Plan',
                          '% customers chruning','% customers continuing'))
colnames(intPlanDf) <- c('a','b')

gIntPlan <- ggplot(intPlanDf,aes(intPlanDf$b,intPlanDf$a))+
        geom_bar(stat = 'identity')+
        labs(title ='Percentage Churn By International Plan', x = 'Customers with International Plan', y = 'Percentage of customers')



# Customer Service calls

ChurnCCC <- sum(churnYes$number_customer_service_calls)/length(churnYes$churn) 
ChurnNoCCC <- sum(churnNo$number_customer_service_calls)/length(churnNo$churn)
dfCCC <- data.frame(c(ChurnCCC,ChurnNoCCC),c('Churning Customers','Continuing Customer'))

gChurnCCC <- ggplot(dfCCC,aes(dfCCC[,2],dfCCC[,1]))+
        geom_bar(stat = 'identity')+
        labs(title ='Average Customer Service calls',
             x = 'Customers', y = 'Average Customer Service Calls')

#Voice Mail Messages

churnVM <- sum(churnYes$number_vmail_messages)/
        length(churnYes[churnYes$voice_mail_plan == 'yes',5])
churnNoVM <- sum(churnNo$number_vmail_messages)/
        length(churnNo[churnNo$voice_mail_plan == 'yes',5])

dfVM <- data.frame(c(churnVM,churnNoVM),c('Churning Customers','Continuing Customer'))

gChurnVM <- ggplot(dfVM,aes(dfVM[,2],dfVM[,1]))+
        geom_bar(stat = 'identity')+
        labs(title ='Average Number of Voice Mail Messages ',
             x = 'Customers', y = 'Number of Voice Mails')


#earnings from customers
earnings <- sum(churnTot$total_day_charge+churnTot$total_eve_charge+churnTot$total_night_charge
                +churnTot$total_intl_charge)
earningsPerCust <- earnings/length(churnTot$churn)

#international charges from customers
 intEarnAvg <- sum(churnTot$total_intl_charge)/length(churnTot$churn)
 intEarnNoAvg <- sum(churnNo$total_intl_charge)/length(churnNo$churn)
 intEarnChurnAvg <- sum(churnYes$total_intl_charge)/length(churnYes$churn)

 #day charges
 dayEarnAvg <- sum(churnTot$total_day_charge)/length(churnTot$churn)
 dayEarnNoAvg <- sum(churnNo$total_day_charge)/length(churnNo$churn)
 dayEarnChurnAvg <- sum(churnYes$total_day_charge)/length(churnYes$churn)
 
 #evening charges
 eveEarnAvg <- sum(churnTot$total_eve_charge)/length(churnTot$churn)
 eveEarnNoAvg <- sum(churnNo$total_eve_charge)/length(churnNo$churn)
 eveEarnChurnAvg <- sum(churnYes$total_eve_charge)/length(churnYes$churn)

 #night charges
 nightEarnAvg <- sum(churnTot$total_night_charge)/length(churnTot$churn)
 nightEarnNoAvg <- sum(churnNo$total_night_charge)/length(churnNo$churn)
 nightEarnChurnAvg <- sum(churnYes$total_night_charge)/length(churnYes$churn)
 
#day minutes 
 dayMntsAvg <- sum(churnTot$total_day_minutes)/length(churnTot$churn)
 dayMntsNoAvg<- sum(churnNo$total_day_minutes)/length(churnNo$churn)
 daysMntsChurnAvg <- sum(churnYes$total_day_minutes)/length(churnYes$churn)

#international minutes

 intMntsAvg <- sum(churnTot$total_intl_minutes)/length(churnTot$churn)
 intMntsNoAvg<- sum(churnNo$total_intl_minutes)/length(churnNo$churn)
 intsMntsChurnAvg <- sum(churnYes$total_intl_minutes)/length(churnYes$churn)

```
The dataset provided contains 19 Predictors and a dependent variable that tells 
whether the customer would churn or not

The categorical predictors are:
State of the customer,
the area code internationalplan,
voice mail plan

The nummerical predictors are:
account_length, area_cod,
number_vmail_messages        
total_day_minutes,
total_day_calls,
total_day_charge,
total_eve_minutes,
total_eve_calls,
total_eve_charge,
total_night_minutes,
total_night_calls,
total_night_charge,
total_intl_minutes,
total_intl_calls,
total_intl_charge,
number_customer_service_calls,
 

Each record contains customer level details.

## The overall churn rate is `r churnRate`
```{r}
#churn Rate
pieChurn <- pie(slices,labels = lbls, col=rainbow(length(lbls)),
  	    main="Percentage Customer chruning")

```
```{r}
print(gIntPlan)
```

```{r}
print(gChurnCCC)
```
We note that on churning customers have made almost twice the number of calls 
to customer service
```{r}
print(gChurnVM)
```
The diffrence in the number of voice mail messages of churning and continuing 
customers is almost negligible

```{r}
print(gChurnstatePerc)
```
We note that California, New Jersey, Texas and Washington have the highest Churn
percentage

#Creating a Logistic Regression MOdel to predict customers who are going to churn

```{r}
#Convert the levels of factors from string to numeric

train$state <- as.numeric(train$state)
train$state <- as.factor(train$state)

test$state <- as.numeric(test$state)
test$state <- as.factor(test$state)

#New levels area_code_408 = 1, area_code_415 =2 area_code_510 =3
train$area_code <- as.numeric(train$area_code )
train$area_code <- as.factor(train$area_code)

test$area_code <- as.numeric(test$area_code )
test$area_code <- as.factor(test$area_code)

#new levels 'yes' = 1, 'no' = 0
train$international_plan <- as.factor(ifelse(train$international_plan == 'yes',1,0))
train$voice_mail_plan <- as.factor(ifelse(train$voice_mail_plan == 'yes',1,0))
train$churn <- ifelse <- as.factor(ifelse(train$churn == 'yes',1,0))

test$international_plan <- as.factor(ifelse(test$international_plan == 'yes',1,0))
test$voice_mail_plan <- as.factor(ifelse(test$voice_mail_plan == 'yes',1,0))

test$churn <- ifelse <- as.factor(ifelse(test$churn == 'yes',1,0))

#creating Logistic Model

logModel <- glm(churn ~ ., family = binomial(),
                        data = train, model = 'TRUE')

# We give 1 to the probability of 0.4 and above to compensate for class imbalance

predictLogProb <- predict(logModel,test[,-20],type = 'response')
predictLogClass  <- as.factor(ifelse( predictLogProb >= 0.4,'1','0'))
logCnm <- confusionMatrix(data = predictLogClass,
                               reference = test$churn)
print(logCnm$overall[1])

print(logCnm$table)

```
#Observations:

From the Logistic Regression model we created above we the get the significance 
Values of the predictors. The predictors which are resposible in determing 'churn'
are :

Inernational Plan : According to our model the customers with an International Plans
are more likely to churn. This suggests that perhaps our international plan is not 
as competitive enough.

Total internation calls: As a corollary to the above observation, the customers who 
use our serivce to make international calls are also more likely to move away. This
confirms our assumption that our International Plan is failing us.

Number of customer service calls : The customer who make contact our customer service
more also ten to churn more. This may suggest that either they are facing issues 
repeatedly or their serivce/request are not being fullfilled by our representatives,
Which is causing them to leave our service

Voice Mail Plan: Our customers who have voice mail plan are also more likely to churn

Number Of Voice Mail Messages: Our customers who get more voice mails seem to churn more
This may suggest network availability issues or that those customers have some other 
provider too, which makes them less available on our service.

#Suggestions
Based on these observations I would suggest the following strategies

Our international plan needs a complete overhaul. We need to analyze the plans
offered by our competitors and come up with an extremly competitive and 
attractive one.

We need to survey our customer service representatives regarding the services 
and requests most made by the customers and introduce any changes if they seem 
feasible. We can further train our customer service representative so that they 
are able to solve our customer's issues in one go.

Or customers with more number of voice mail messages also seem to churn more, 
we can survey our customers to understand whether there are connectivity issues
at some places and rectify these issues if they exist.

#Creating a gradient boosting model using XGBoost
```{r}

#function to create boosted trees.
createBT <- function(trainData,testData,n){
        
        target <- as.numeric(levels(trainData$churn))[trainData$churn]
        boosTree <- xgboost(data = data.matrix(trainData[,-20]), label = target,
                            max.depth = 6,nrounds = n,verbose = 0,
                            objective = 'binary:logistic',eta = 0.3)


        predictBTProb <- predict(boosTree,data.matrix(testData[,-20]),
                                 type = 'response')
        predictBTClass <- as.factor(ifelse( predictBTProb >= 0.5,1,0))

        btCnm <- confusionMatrix(data = predictBTClass,
                               reference = testData$churn)

        acc <- (btCnm$table)
        return(btCnm$overall[1])
}
        

#3 Fold Cross Validation
set.seed(111)
cv_train_index <- createFolds(seq_len(nrow(train)), k = 3, list = TRUE,
                                                             returnTrain = TRUE)
#Tuning the model
avgAcc <- NA
acc <- NA
k <- 1
for(i in seq(from = 1, to = 21 , by = 2)){
        acc <- NA
        
        for(j in 1:3){
                
                cvTrain <- train[cv_train_index[[j]],]
                cvTest <- train[-cv_train_index[[j]],]
                
                acc[j] <- createBT(cvTrain,cvTest,i)
                cnm <- NA
        
        }
         avgAcc[k] <- mean(acc)
         k <- k+1
}

ledger_mat <- data.frame(seq(from = 1, to = 21 , by = 2),avgAcc)

#plotting Accuracy Vs. Number of boosting rounds
ggplot(ledger_mat, aes(ledger_mat[,1],ledger_mat[,2]))+
geom_point()+ geom_line()+xlab('No of Rounds')+ylab('Accuracy')


bestRounds <- ledger_mat[which.max(ledger_mat[,2]),1]
bestAcc <-  ledger_mat[which.max(ledger_mat[,2]),2]       

#build the final model using the tuned number of rounds

target <- as.numeric(levels(train$churn))[train$churn]
finalBoosTree <- xgboost(data = data.matrix(train[,-20]), label = target,
                         max.depth = 6,nrounds = bestRounds,verbose = 0,
                         objective = 'binary:logistic',eta = 0.3)


predictBTProb <- predict(finalBoosTree,data.matrix(test[,-20]), type = 'response')
predictBTClass <- as.factor(ifelse( predictBTProb >= 0.5,1,0))

btCnm <- confusionMatrix(data = predictBTClass,
                               reference = test$churn)

print(btCnm$overall[1])
print(btCnm$table)

featureImp <- xgb.importance( feature_names = names(train),model = finalBoosTree)
print(featureImp)

prefBT <- performance(prediction(predictBTProb,test$churn),
                  measure = "tpr",x.measure = "fpr")
plot(prefBT)

#auc
aucBT <- performance(prediction(predictBTProb,test$churn),measure = "auc")
print(c("Area Under the Curve: ", aucBT@y.values[[1]]),quote = FALSE)

```

#Observations
 *Average day charges  incurred by all the customers = 30.64967
 *Average daytime minutes consumed by all customers =  180.2889
 
 *Average day charges incurred by the customers who will not churn =   29.87749
 *Average daytime minutes consumed by customers who will not churn =   175.7466
 
 *Average day charges incurred by cuatomers who churn = 35.33842
 *Average daytime minutes consumed by customers who churn = 207.8706
 
 *It appears that our customers who churn tend to call more during daytime and 
 *are obviously unsatistfied by our rates hence they churn.
 
 *Our current daytime rate is : $0.17 per minute.
 
 58% of our customers who have international plan churn.
 So wee need to do something about this too.
 
```{r}
 print("Taking a look at the confusion matrix")
 print(btCnm$table)
```
### No. of churning customers predicted = 163
### No. of churning customers not predicted = 61
### No. of continuing customers wrongly predicted as churning = 61
 
 
 
 My plan is to give the customers predicted as churning 
 300 mintues of free daytime calling and 20$ of Google Voice amount  
 for a  contract of 6 months
 
#Assumptions:
 
 Monthy incomming earnings from a customer        = $60.44
 Monthly Cost of providing service to a customer  = .925*60.44 = $55.907
 (92.5% of incoming cost)
 
 Monthly earninngs form a single customer         = $60.44-$55.907 = 4.533
 Targeted Customers wont churn for 6 months due to contract
 
#Note:
 
 Average day minutes consumed by targeted customer per month     = 208
 Average evening minutes consumed by targeted customer per month = 210
 Average day minutes consumed by targeted customer per month      = 207
 
 
 Our investment per targeted customer is 20$ for 6 months and 300 daytime call
 minutes.
 
 We assume that customer will use up 208 free daytime minutes in the first month
 and the remaining 92 minutes in the second month.
 We assume that the customer uses the google voice amount provided for 
 international calls
 
 Total no of customers we targeted are 171 but the calculations below are for the
 163 customer who were surely going to churn.
 
#Calculations 
 1st Month cost for targeted customer:
 daytime call cost = 0*0.17   = $0.00
 evening call cost = 210*0.08 = $16.8
 night call cost =   207*0.04 = $8.04

 Total incoming from targeted customer in 1st month = 16.8+8.28 = $25.08
 Operating costs                                                = $23.19
 Total earnings from 1 customer for 1st month  = 25.08-23.19    = $1.89
 No of targeted Customers = 163
 Total earning for 1st month = 163*1.89                         = $308.07
 
 
 2nd Month cost for targeted customer:
 daytime call cost = 92*0.17  = $15.64
 evening call cost = 210*0.08 = $16.8
 night call cost =   207*0.04 = $8.28
 
 Total incoming from targeted customer in 2nd month = 16.8+8.28 = $40.72
 Operating costs                                                = $37.66
 Total earnings from 1 customer for 3rd month  = 25.08-23.19    = $3.05
 No of targeted Customers = 163
 Total earning for 1st month = 163*3.05                        = $495.17
 
 
 3rd Month cost for targeted customer:
 daytime call cost = 208*0.17 = $35.56
 evening call cost = 210*0.08 = $16.8
 night call cost =   207*0.04 = $8.04

 Total incoming from targeted customer in 1st month = 16.8+8.28 = $60.44
 Operating costs                                                = $55.90
 Total earnings from 1 customer for 1st month  = 25.08-23.19    = $4.533
 No of targeted Customers = 163
 Total earning for 3rd month = 163*4.533                        = $738.879
 
 
 For the 4th, 5th and 6th month earnings from each customer is the same as that 
 of 3rd month. 
 Earning for 4th month  = $738.879
 Earning for 5th month  = $738.879
 Earning for 6th month  = $738.879
 
 
###Monthly Dollar value gained by retaining a customer is $4.533 
 
 Total revenue lost on targeting the 8 customers who were not going to
 churn = $68.52
 
 Total earnings from targeted customer who were going to churn at the end of 5
 months is $3021.72.
 
 Total earnings from targeted customer who were going to churn at the end of 6
 months is $3760.00
 
### Total revenue lost on targeting the 8 customers who were not going to churn = $68.52.
 
 Total investment is $20*163 = $3420..The 300 free mins are included in calculations.
 We are guaranteed to meet our breakeven point by the end of sixth month
 
 As I was able to predict 163 out of the 224 customers who were going to churn
 and was able to comeup with a plan that is profiatable, I conclude that the 
 model is good 
 
